Finding Maximum $k$-Cliques Faster using Lazy Global Domination

### 2016-04-05 paper/paper.tex
All hail eggy

Here we give experimental results on a range of standard benchmarks, and on real-world and random
graphs. Where timing results are reported, the experiments were run on a machine with Intel E5645
processors, and single-threaded runtimes are given. The time taken to read in the graph from a file
is excluded, but preprocessing time (including the construction of $G^k$ and the bitset encoding) is
included. We use the term \emph{nodes} to refer to the number of recursive calls made by the
branch-and-bound part of the algorithm.

@@

In \cref{table:erdos} we present experimental results from Erd\H{o}s
collaboration graphs from the Pajek dataset by Vladimir Batagelj and Andrej
Mrvar\footnote{http://vlado.fmf.uni-lj.si/pub/networks/data/}.  We were able to solve all of
these problems in under eight minutes (and all but three in under two seconds) when using the
domination rule. However, using the unmodified maximum clique algorithm, two of these results did
not finish running within one day. Note that for $k = 4$, a $k$-clique covers all of ``Erdos02''.

@@

With the domination rule, we solve all of these problems within a tenth of a second. Without, two of
the results take over a day, and the rest remain trivial. Note that in several cases, for some
values of $k$ a $k$-clique covers the entire graph.  Again using Chang et al.'s results
\cite{Chang:2013}, we see that for the first six graphs in this table the $k$-clique and $k$-club
numbers are the same for each value of $k$ (Chang et al.\ did not investigate ``c-fat500-10'' or
``p-hat300-1'').

@@

\Cref{table:partitioning} presents results from the smallest 20 partitioning graphs
from the 10th DIMACS Implementation
Challenge\footnote{http://staffweb.cms.gre.ac.uk/\textasciitilde{}wc06/partition/}.  Many of these graphs are
considerably larger than those typically considered for the maximum clique problem, and we might
expect our $O(|V|^2)$ memory requirements to cause problems. Nonetheless, with the domination rule
there is only one instance which we were unable to solve within a day (and without the domination
rule, there are two).

@@

\Cref{table:clustering} presents results from the smallest 20 clustering graphs
from the 10th DIMACS Implementation
Challenge\footnote{http://www.cc.gatech.edu/dimacs10/archive/clustering.shtml}. Again, from a
maximum clique perspective these would be considered unusually large graphs. However, only five were
unsolvable within a day (plus a sixth when the domination rule was not used), and half of the
problems took under two seconds.

>>>>

Here we give experimental results on a range of standard benchmarks, and on real-world and random
graphs. Where timing results are reported, the experiments were run on a machine with Intel E5-2687W
v3 processors, and single-threaded runtimes are given. The time taken to read in the graph from a
file is excluded, but preprocessing time (including the construction of $G^k$ and the bitset
encoding) is included. We use the term \emph{nodes} to refer to the number of recursive calls made
by the branch-and-bound part of the algorithm.

@@

In \cref{table:erdos} we present experimental results from Erd\H{o}s
collaboration graphs from the Pajek dataset by Vladimir Batagelj and Andrej
Mrvar\footnote{http://vlado.fmf.uni-lj.si/pub/networks/data/}.  We were able to solve all of
these problems in under eight minutes (and all but three in under two seconds) when using the
domination rule. However, using the unmodified maximum clique algorithm, two of these results did
not finish running within one hour. Note that for $k = 4$, a $k$-clique covers all of ``Erdos02''.

@@

With the domination rule, we solve all of these problems within a tenth of a second. Without, two of
the results take over an hour, and the rest remain trivial. Note that in several cases, for some
values of $k$ a $k$-clique covers the entire graph.  Again using Chang et al.'s results
\cite{Chang:2013}, we see that for the first six graphs in this table the $k$-clique and $k$-club
numbers are the same for each value of $k$ (Chang et al.\ did not investigate ``c-fat500-10'' or
``p-hat300-1'').

@@

\Cref{table:partitioning} presents results from the smallest 20 partitioning graphs
from the 10th DIMACS Implementation
Challenge\footnote{http://staffweb.cms.gre.ac.uk/\textasciitilde{}wc06/partition/}.  Many of these graphs are
considerably larger than those typically considered for the maximum clique problem, and we might
expect our $O(|V|^2)$ memory requirements to cause problems. Nonetheless, with the domination rule
there is only one instance which we were unable to solve within an hour (and without the domination
rule, there are two).

@@

\Cref{table:clustering} presents results from the smallest 20 clustering graphs
from the 10th DIMACS Implementation
Challenge\footnote{http://www.cc.gatech.edu/dimacs10/archive/clustering.shtml}. Again, from a
maximum clique perspective these would be considered unusually large graphs. However, only five were
unsolvable within an hour (plus a sixth when the domination rule was not used), and half of the
problems took under two seconds.
